---
title: 'R Course: Lesson 5: ANALYSIS OF VARIANCE (ANOVA)'
author: "Reddy Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(gapminder)
library(forcats)
library(stargazer)
library(purrr)
```

# Math
ANOVA is used only in continuous dependent variables

## example in R

```{r}
gapdata <- gapminder %>% 
  filter(year==2007&
           continent %in% c("Americas","Europe","Asia")) %>% 
  select(continent,lifeExp)

gapdata %>% 
  group_by(continent) %>% 
  summarise(mean.life=mean(lifeExp)) %>% 
  arrange(mean.life) %>% 
  ungroup()

gapdata %>% 
  aov(lifeExp~continent,data=.) %>% 
  summary()

gapdata %>% 
  aov(lifeExp~continent,data=.) %>% 
  TukeyHSD() %>% 
  plot()
```
```{r}
my.data <- msleep %>% 
  select(vore,sleep_rem) %>% 
  drop_na()

my.data %>% 
  aov(sleep_rem~vore,data=.) %>% 
  TukeyHSD() %>% 
  plot()

model.1 <- aov(sleep_rem~vore,data=my.data) 

summary(model.1)
```
```{r}
model.1 <- aov(sleep_rem~vore,data=my.data) 

summary(model.1)
my.data %>% 
  aov(sleep_rem~vore,data=.) %>% 
  TukeyHSD()
```
```{r}
library(palmerpenguins)
table(penguins$species)
penguins %>% 
  ggplot(aes(x=species,y=flipper_length_mm))+
  geom_boxplot()
  
```
```{r}
penguins %>% 
  aov(flipper_length_mm~species,data=.) %>% 
  TukeyHSD()
```
```{r}
summary(penguins %>% 
  aov(flipper_length_mm~species,data=.))
```
```{r}
penguins %>% 
  ggplot(aes(x=flipper_length_mm))+
  geom_histogram()+
  facet_wrap(~species,ncol=1)
```

```{r}
penguins %>% 
  group_by(species) %>% 
  summarise(var(flipper_length_mm,na.rm = T))
```
```{r}
model.2 <- aov(flipper_length_mm~species,data=penguins)
summary(model.2)
TukeyHSD(model.2)
```
# Data cleaninng
```{r}
data.election.results <- list.files(path="data5/elections",full.names = T) %>% 
  map(read.table,header=T,sep="\t") %>% 
  reduce(rbind)

data.election <- read.csv("data5/rcourse_lesson5_data_elections.txt",sep="\t")
data.states <- read.csv("data5/rcourse_lesson5_data_states.txt",sep="\t")
xtabs(~civil_war,data.states)
xtabs(~incumbent_party,data.election)
```
```{r}
data.states.clean <- data.states %>% 
  filter(!is.na(civil_war)) %>% 
  group_by(civil_war) %>% 
  arrange(order_enter) %>% 
  filter(row_number()<=11) %>% 
  ungroup()

xtabs(~civil_war,data.states.clean)

data.clean <- data.election.results %>% 
  inner_join(data.election) %>% 
  inner_join(data.states.clean) %>% 
  mutate(state=factor(state))

xtabs(~incumbent_party+civil_war,data = data.clean)
```
# Figures
```{r}
data.figs <- data.clean %>% 
  mutate(civil_war=factor(civil_war,levels=c("union","confederacy"),labels=c("Union","Confederacy"))) %>% 
  mutate(incumbent_party=factor(incumbent_party,levels=c("democrat","republican"),labels=c("Democrat","Republican")))

incumbent.histogram.full.plot <- data.figs %>% 
  ggplot(aes(x=perc_votes_incumbent,fill=incumbent_party))+
  geom_histogram(bins = 30)+
  facet_grid(incumbent_party~civil_war)+
  scale_fill_manual(values=c("blue","red"))

incumbent.histogram.full.plot
```
```{r}
data.figs.state.sum <- data.figs %>% 
  group_by(state,incumbent_party,civil_war) %>% 
  summarise(perc_incumbent_mean=mean(perc_votes_incumbent,na.rm = T)) %>% 
  ungroup()

incumbent.histogram.full.plot1 <- data.figs.state.sum %>% 
  ggplot(aes(x=perc_incumbent_mean,fill=incumbent_party))+
  geom_histogram(bins = 30)+
  facet_grid(incumbent_party~civil_war)+
  scale_fill_manual(values=c("blue","red"))

incumbent.histogram.full.plot1
```
```{r}
incumbent.boxplot <- data.figs.state.sum %>% 
  ggplot(aes(x=civil_war,y=perc_incumbent_mean,fill=incumbent_party))+
  geom_boxplot()+
  ylim(0,100)+
  geom_hline(yintercept = 50)+
  scale_fill_manual(values = c("blue","red"))

incumbent.boxplot
```
```{r}
data.figs.sum <- data.figs.state.sum %>% 
  group_by(incumbent_party,civil_war) %>% 
  summarise(mean=mean(perc_incumbent_mean,na.rm = T),sd=sd(perc_incumbent_mean,na.rm=T),n=n()) %>% 
  ungroup() %>% 
  mutate(se=sd/sqrt(n),se_high=mean+se,se_low=mean-se)
  
barplot.plot <- data.figs.sum %>% 
  ggplot(aes(x=civil_war,y=mean,fill=incumbent_party))+
  geom_bar(stat = "identity",position="dodge")+
  geom_errorbar(aes(ymin=se_low,ymax=se_high),width=.2,position=position_dodge(.9))+
  ylim(0,100)+
  geom_hline(yintercept = 50)+
  scale_fill_manual(values=c("blue","red"))+
  ggtitle("How we vote")+
  theme(text=element_text(size=14),title=element_text(size=16))+
  ggeasy::easy_center_title()
  
barplot.plot
```


# Statistics
```{r}
install.packages("ez")
library(ez)

data.stats <- data.clean %>% 
  mutate(civil_war=factor(civil_war,levels=c("union","confederacy"))) %>% 
  group_by(state,incumbent_party,civil_war) %>% 
  summarise(perc_incumbent_mean=mean(perc_votes_incumbent,na.rm = T)) %>% 
  ungroup()

xtabs(~state+incumbent_party,data.stats)
xtabs(~state+civil_war,data.stats)
```
```{r}
incumbent.aov <- aov(perc_incumbent_mean~incumbent_party*civil_war+Error(state/incumbent_party),data = data.stats)

summary( incumbent.aov)
#TukeyHSD(incumbent.aov)
```
## ezANOVA
```{r}
incumbent.ezanova <- ezANOVA(data.frame(data.stats),dv=perc_incumbent_mean,
                             wid=state,within=incumbent_party,between=civil_war,type=3)

summary(incumbent.ezanova)
```
```{r}
data.union.stats <- data.stats %>% 
  filter(civil_war=="union") %>% 
  spread(incumbent_party,perc_incumbent_mean)

data.confederacy.stats <- data.stats %>% 
  filter(civil_war=="confederacy") %>% 
  spread(incumbent_party,perc_incumbent_mean)

data.democrat.stats <- data.stats %>% 
  filter(incumbent_party=="democrat")
data.republican.stats <- data.stats %>% 
  filter(incumbent_party=="republican")

```

## t test: Effect of incumbent party
```{r}
incumbent.union.ttest <- t.test(data.union.stats$democrat,data.union.stats$republican,paired = T)
incumbent.union.ttest

incumbent.confederacy.ttest <- t.test(data.confederacy.stats$democrat,data.confederacy.stats$republican,paired = T)
incumbent.confederacy.ttest
```
```{r}
incumbent.democrat.ttest <- t.test(perc_incumbent_mean~civil_war,paired=F,data = data.democrat.stats)


incumbent.republican.ttest <- t.test(perc_incumbent_mean~civil_war,paired=F,data = data.republican.stats)

incumbent.democrat.ttest
incumbent.republican.ttest
```


## Incumbent Party: Do Democrats or Republicans get a higher percentage of the vote when they are an incumbent?

## Civil War Country: Do Union or Confederate states vote differently for incumbents?

## Incumbent Party x Civil War Country: Is there an interaction between these variables?